---
title: "GAM_BETA_SD_Prior_Simulation"
author: "Adam C. Smith"
date: "08/03/2022"
output: pdf_document
bibliography: references.bib
---

```{r setup, message=FALSE,echo=FALSE}
options(scipen=99999)
library(tidyverse)
library(cmdstanr)
library(mgcv) 
library(patchwork)
library(kableExtra)

```

# Prior Simulation of the SD of GAM parameters

Running a prior simulation for alternative prior distributions on the SD of the GAM BETA parameters, which are the hyperparameters that control the shape and wiggliness of the survey-wide mean smoothed population trajectory. This simulation compares the effects of alternative priors on derived estimates of the long-term (53-year), and short-term (10-year) population trend estimates that would result from a population trajectory estimated with the same spline basis function used in this paper. This SD parameter controls the scale and variation of the parameters that link the spline basis function to the estimated smooth population trajectory. The posterior estimate of this SD parameter is what determines the complexity (wiggliness) and magnitude of the population change ([@crainiceanu2005; @bürkner2017; @goodrich2020]).

The semi-parametric nature of smooths and the variation among alternative basis functions complicates the use of default or informative priors ([@lemoine2019; @banner2020]). We've used a prior simulation to translate these alternative priors into intuitive values of population trends that are directly interpretable as biological parameters. We then compare the prior distribution of trends that would result from these alternative priors to:

1.  the collection of realised long-term trend estimates from a different statistical model applied to the North American Breeding Bird Survey ([@link2020]).

2.  our own prior knowledge about probable rates of change in wild bird populations at continental scales.

We compared half normal and half t-distributions for the priors on the Standard Deviation on the GAM parameters.

1.  normal

2.  t-distribution with 3 degrees of freedom

And, for each of the half-normal and half-t-distributions, we compared 5 different values to set the prior-scale: (0.5, 1, 2, 3, and 4). Given the log-link in the trend model and the scaling of the low-rank thin-plate regression spline with identifiability constraints ([@wood2020]), these 5-values of prior-scales should cover the range of plausible parameter values.

Here's the code that runs the simulations in Stan.

```{r run_simulations,eval=FALSE}

# this is all set-up
library(tidyverse)
library(cmdstanr)
library(mgcv) 
library(patchwork)


for(pp in c("norm","t")){
  for(prior_scale in c(0.5,1,2,3,4)){
    
  tp = paste0("GAM_",pp,prior_scale,"_rate")

  #STRATA_True <- log(2)
  output_dir <- "output/"
  out_base <- tp
  csv_files <- paste0(out_base,"-",1:3,".csv")
  

  if(pp == "norm"){
    pnorm <- 1
  }
  if(pp == "t"){
    pnorm <- 2
  }
  
  if(!file.exists(paste0(output_dir,csv_files[1]))){
    
    nyears = 54 #to match the time-scales of BBS and CBC analyses
  dat = data.frame(year = 1:nyears)
  nknots = 13  
  nknots_realised = nknots-1 
  
  M = mgcv::smoothCon(s(year,k = nknots, bs = "tp"),data = dat,
                       absorb.cons=TRUE,#this drops the constant
                       diagonal.penalty=TRUE) ## If TRUE then the smooth is reparameterized to turn the penalty into an identity matrix, with the final diagonal elements zeroed (corresponding to the penalty nullspace). This fits with the Bayesian interpretation of the complexity penalty as the inverse of the variance of the i.i.d. collection of parameters.
    
    
    year_basis = M[[1]]$X
    
    
    
    
    stan_data = list(#scalar indicators
      nyears = nyears,

      
      #GAM structure
      nknots_year = nknots_realised,
      year_basis = year_basis,
      
      prior_scale = prior_scale,
      pnorm = pnorm
      )
    
    
    
    
    # Fit model ---------------------------------------------------------------
    
    print(paste("beginning",tp,Sys.time()))
    
    mod.file = "models/GAM_prior_sim.stan"
    
    ## compile model
    model <- cmdstan_model(mod.file)
    
    
    # Initial Values ----------------------------------------------------------
    
    
    init_def <- function(){ 
      list(sdbeta = runif(1,0.01,0.1),
           BETA_raw = rnorm(nknots_realised,0,0.01))}
    
    stanfit <- model$sample(
      data=stan_data,
      refresh=100,
      chains=2, iter_sampling=1000,
      iter_warmup=500,
      parallel_chains = 2,
      #pars = parms,
      adapt_delta = 0.8,
      max_treedepth = 14,
      seed = 123,
      init = init_def,
      output_dir = output_dir,
      output_basename = out_base)
    
    
   save(list = c("stanfit","stan_data","csv_files",
                  "out_base"),
         file = paste0(output_dir,"/",out_base,"_gamye_iCAR.RData"))
    
    
    
  }
  
}#end prior_scale loop
}#end pp loop



```

## The simulation model

The Stan simulation model is very simple. We implemented it in Stan to match the implementation in the full model (although this simulation doesn't require MCMC sampling).

```{stan, output.var = "GAM_prior_sim.stan", eval = FALSE}
// simple GAM prior simulation

data {
  int<lower=1> nyears;
  //scale of the prior distribution
  real<lower=0>  prior_scale; 
  // indicator for the prior distribution 0 = t, 1 = normal
  int<lower=0,upper=1> pnorm; 
// data for spline s(year)
  // number of knots in the basis function for year
  int<lower=1> nknots_year;  
  // basis function matrix
  matrix[nyears, nknots_year] year_basis; 
}

parameters {
  real<lower=0> sdbeta;    // sd of spline coefficients 
  // unscaled spline coefficients
  vector[nknots_year] BETA_raw;         
}
 
transformed parameters { 
  vector[nknots_year] BETA;  //scaled spline parameters       
  vector[nyears] smooth_pred;

    BETA = sdbeta * BETA_raw; //scaling the spline parameters
     smooth_pred = year_basis * BETA; //log-scale smooth trajectory
  }
  
model {

//Conditional statements to select the prior distribution
if(pnorm == 1){
 sdbeta ~ normal(0,prior_scale); //prior on sd of GAM parameter 
}
if(pnorm == 0){
  sdbeta ~ student_t(3,0,prior_scale); //prior on sd of GAM parameter
}

   BETA_raw ~ normal(0,1); //non-centered parameterisation

}

 generated quantities {
  //estimated smooth on a count-scale
   vector[nyears] nsmooth = exp(smooth_pred);
    
  }


```

We then summarized the estimated trajectories as well as all possible 1-year, 10-year, and long-term trends from the alternative priors.

```{r summarising,eval=FALSE}

source("Functions/posterior_summary_functions.R") ## loads the saved simulation results

nsmooth_out <- NULL
trends_out <- NULL
summ_out <- NULL

for(pp in c("norm","t")){
  for(prior_scale in c(0.5,1,2,3,4)){
    
    tp = paste0("GAM_",pp,prior_scale,"_rate")
    
    #STRATA_True <- log(2)
    output_dir <- "output/"
    out_base <- tp
   

load(paste0(output_dir,"/",out_base,"_gamye_iCAR.RData"))

summ = stanfit$summary()

summ <- summ %>% 
  mutate(prior_scale = prior_scale,
         distribution = pp)

nsmooth_samples <- posterior_samples(stanfit,
                                 parm = "nsmooth",
                                 dims = c("Year_Index"))



BETA_samples <- posterior_samples(stanfit,
                                  parm = "BETA",
                                  dims = c("k"))
BETA_wide <- BETA_samples %>% 
  pivot_wider(.,id_cols = .draw,
              names_from = k,
              names_prefix = "BETA",
              values_from = .value)


nsmooth_samples <- nsmooth_samples %>% 
  left_join(., BETA_wide,by = ".draw") %>% 
  mutate(prior_scale = prior_scale,
         distribution = pp)



nyears = max(nsmooth_samples$Year_Index)
# function to calculate a %/year trend from a count-scale trajectory
trs <- function(y1,y2,ny){
  tt <- (((y2/y1)^(1/ny))-1)*100
}

for(tl in c(2,11,nyears)){ #estimating all possible 1-year, 10-year, and full trends
  ny = tl-1
  yrs1 <- seq(1,(nyears-ny),by = ny)
  yrs2 <- yrs1+ny
  for(j in 1:length(yrs1)){
    y2 <- yrs2[j]
    y1 <- yrs1[j]
    
nyh2 <- paste0("Y",y2)
nyh1 <- paste0("Y",y1)
trends <- nsmooth_samples %>% 
  filter(Year_Index %in% c(y1,y2)) %>% 
  select(.draw,.value,Year_Index) %>% 
  pivot_wider(.,names_from = Year_Index,
              values_from = .value,
              names_prefix = "Y") %>%
  rename_with(.,~gsub(pattern = nyh2,replacement = "YE", .x)) %>% 
  rename_with(.,~gsub(pattern = nyh1,replacement = "YS", .x)) %>% 
  group_by(.draw) %>% 
  summarise(trend = trs(YS,YE,ny))%>% 
  mutate(prior_scale = prior_scale,
         distribution = pp,
         first_year = y1,
         last_year = y2,
         nyears = ny)
trends_out <- bind_rows(trends_out,trends)
}
}
nsmooth_out <- bind_rows(nsmooth_out,nsmooth_samples)
summ_out <- bind_rows(summ_out,summ)
print(paste(pp,prior_scale))

  }#prior_scale
}# pp

save(file = "output/GAM_prior_sim_summary.RData",
     list = c("nsmooth_out",
              "trends_out",
              "summ_out"))


```



# Comparing simulation priors to realised data

## Realised trend estimates

First, here is the distribution of long-term trends from a different model for the BBS data from 1966-2019, for 426 species.

```{r realised bbs trends, fig.show='asis', fig.dim = c(8, 8)}
bbs_trends_usgs <- read.csv("data/BBS_1966-2019_core_best_trend.csv")


## selecting survey-wide trend estimates
bbs_trends_usgs_long <-bbs_trends_usgs %>% 
  filter(Region == "SU1")%>%  
  select(Trend,Species.Name) %>% 
  mutate(abs_trend = abs(Trend)) %>% #calculating absolute values of the trends
  arrange(-abs_trend)

G_long_usgs <- max(bbs_trends_usgs_long$abs_trend)
  
realised_long_bbs_hist <- ggplot(data = bbs_trends_usgs_long,
                            aes(abs_trend,after_stat(density)))+
  geom_freqpoly(breaks = seq(0,13,0.5),center = 0)+
  xlab("Absolute value of long-term BBS trends USGS models (1966-2019)")+
  theme_bw()+
  scale_y_continuous(limits = c(0,1))
print(realised_long_bbs_hist)

```

The maximum absolute value of an observed long-term trends are for Cave Swallow and Eurasian Collared Dove, which have increased at an annual rate of `r round(G_long_usgs,1)` %/year. This annual rate of change implies an approximate `r signif(((((G_long_usgs/100)+1)^53)-1)*100,2)` % overall increase in the populations since 1966. As such, we feel this represents example of an "extreme" long-term trend that is unlikely to be observed in most of the BBS dataset. For example, the next largest values of trend in the data is a `r round((bbs_trends_usgs_long$abs_trend[3]),1)` %/year increase in Canada Goose populations.

```{r realised trend header}

kable(bbs_trends_usgs_long[1:3,], booktabs = TRUE,
      digits = 3) %>%
  kable_styling(font_size = 8)

```

And the largest absolute value of trends for a declining species is \< 4%/year for King Rail, Bank Swallow, and Lark Bunting.

```{r realised declines header}
bbs_declines_usgs_long <-bbs_trends_usgs_long %>% 
  filter(Trend < 0)%>%  
  arrange(-abs_trend)

kable(bbs_declines_usgs_long[1:3,], booktabs = TRUE,
      digits = 3) %>%
  kable_styling(font_size = 8)

```


## Summarize the long-term trends from the prior simulations

```{r prior_trend_distributions}
 
load("output/GAM_prior_sim_summary.RData")

trends_out <- trends_out %>% 
  mutate(abs_trend = abs(trend),# absolute values of trends
         scale_factor = factor(prior_scale,ordered = TRUE)) #just for plotting

trends_long <- trends_out %>%  #select long-term trends only
  filter(nyears == 53 ) 

trends_normal <- trends_long %>% 
  filter(distribution == "norm")
```
The distributions for the normal priors do a reasonable job of covering the range of possible long-term trend values, but the heavier-tailed t-distributions seem to better fit the shape of the distributions.

```{r overplots long-term}

  overp_normal <- realised_long_bbs_hist +
    geom_freqpoly(data = trends_normal,
                aes(abs_trend,after_stat(density),
                    colour = scale_factor),
                breaks = seq(0,100,1),center = 0,
                alpha = 0.8)+
    scale_colour_viridis_d(begin = 0.5,alpha = 0.8,
                           "Prior Scale\nSD half-normal")+
  coord_cartesian(xlim = c(0,20))
    
  
  trends_t <- trends_long %>% 
  filter(distribution == "t")


  overp_t <- realised_long_bbs_hist +
    geom_freqpoly(data = trends_t,
                aes(abs_trend,after_stat(density),
                    colour = scale_factor),
                breaks = seq(0,100,1),center = 0,
                alpha = 0.8)+
    scale_colour_viridis_d(begin = 0.5,alpha = 0.8,
                           "Prior Scale\nSD half-t df=3")+
  coord_cartesian(xlim = c(0,20))
 
```

The half-t-distributions with a scale value of 2 or 3 fit the shape of the realised trend distribution reasonably well, and the long-tail includes significant prior mass at and beyond the observed maximum absolute values of trends. 

```{r plot_prior_trend_distributions, fig.show='asis', fig.dim = c(8, 10)}
 print(overp_normal / overp_t)
```

In addition, each of the prior distributions include some prior mass at trend values beyond the realised maximum values. And the broadest priors (scales of 3 or 4), include prior mass at trend values that are extremely unlikely for a wild population to sustain over >50 years.

```{r prior_table}
quant_long_tends <- trends_long %>% 
  group_by(distribution,prior_scale) %>% 
  summarise(median_abs_t = median(abs_trend),
            U80 = quantile(abs_trend,0.80),
            U90 = quantile(abs_trend,0.90),
            U99 = quantile(abs_trend,0.99),
            pGTmax = length(which(abs_trend > G_long_usgs))/length(abs_trend))

kable(quant_long_tends, booktabs = TRUE,
      digits = 3) %>%
  kable_styling(font_size = 8)

 
```


# References
